---
title: "Benchmark"
author: "Michel Lang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r,include=FALSE}
do.eval = requireNamespace("microbenchmark", quietly = TRUE)
```

This small benchmark compares the performance of the base64 encoding/decoding in package `base64url` with the implementations in the packages [`base64enc`](https://cran.r-project.org/package=base64enc) and [`openssl`](https://cran.r-project.org/package=openssl).

## Encoding of a single string
```{r, eval=do.eval}
library(base64url)
library(base64enc)
library(openssl)
library(microbenchmark)

x = "plain text"
microbenchmark(
  base64url = base64_urlencode(x),
  base64enc = base64encode(charToRaw(x)),
  openssl = base64_encode(x)
)
```

## Decoding of a single string
```{r, eval = do.eval}
x = "N0JBLlRaUTp1bi5KOW4xWStNWEJoLHRQaDZ3"
microbenchmark(
  base64url = base64_urldecode(x),
  base64enc = rawToChar(base64decode(x)),
  openssl = rawToChar(base64_decode(x))
)
```

## Encoding and decoding of character vectors
Here, the task has changed from encoding/decoding a single string to processing multiple strings stored inside a character vector.
First, we create a small utility function which returns `n` random strings with a random number of characters (between 1 and 32) each.
```{r, eval = do.eval}
rand = function(n, min = 1, max = 32) {
  chars = c(letters, LETTERS, as.character(0:9), c(".", ":", ",", "+", "-", "*", "/"))
  replicate(n, paste0(sample(chars, sample(min:max, 1), replace = TRUE), collapse = ""))
}
set.seed(1)
rand(10)
```

Only `base64url` is vectorized for string input, the alternative implementations need wrappers to process character vectors:
```{r, eval = do.eval}
base64enc_encode = function(x) {
  vapply(x, function(x) base64encode(charToRaw(x)), NA_character_, USE.NAMES = FALSE)
}

openssl_encode = function(x) {
  vapply(x, function(x) base64_encode(x), NA_character_, USE.NAMES = FALSE)
}

base64enc_decode = function(x) {
  vapply(x, function(x) rawToChar(base64decode(x)), NA_character_, USE.NAMES = FALSE)
}

openssl_decode = function(x) {
  vapply(x, function(x) rawToChar(base64_decode(x)), NA_character_, USE.NAMES = FALSE)
}
```

The following benchmark measures the runtime to encode 1000 random strings and then decode them again:
```{r, eval = do.eval}
set.seed(1)
x = rand(1000)
microbenchmark(
  base64url = base64_urldecode(base64_urlencode(x)),
  base64enc = base64enc_decode(base64enc_encode(x)),
  openssl = openssl_decode(openssl_encode(x))
)
```
